---
title: "R Notebook"
output: html_notebook
---


Importing and Exploratory Data Analysis

```{r}
require("stats")
require("forecast")
require("MLmetrics")

df = read.table("co2_mm_mlo.txt", 
                col.names = c("Year", "Month", "Date", "average", "interpolated", "trend", "days"))
df[,8] = as.Date(paste(df$Year, df$Month, 1, sep = "-"), format = "%Y-%m-%d")
plot(df[,8], df[,5], type = "l", xlab = "Year", ylab = "Yearly Average", 
     main = "Mauna Loa Yearly Carbon Dioxide Concentration")
plot(1958:2018, tapply(df[,5], format(df[,8], "%Y"), mean), xlab = "Year", 
     ylab = "Yearly Average", main = "Mean Annual CO2 Levels", type = "l")
plot(df[647:718, 8], df[647:718, 4], type = "o", xlab = "Years", 
     ylab = "Average CO2 Level (ppm)", main = "Recent Monthly Mean CO2 Levels")
```

This is for a simple prediction on the training data
```{r}
shortts = ts(df[647:718, 4], frequency = 12, start = c(2012, 1))  # only 2012-2017
fullts = ts(df[1:718,5], frequency = 12, start = c(1958, 3))      # 1953 - 2017
plot(decompose(fullts))
plot(HoltWinters(shortts))
plot(HoltWinters(fullts))
MAPE(HoltWinters(shortts)$fitted[,1], df[659:718, 4])
MAPE(HoltWinters(fullts)$fitted[,1], df[13:718,5])
```

This is for a prediction of future seasonal data using Holt-Winters 
```{r}
pred = data.frame(predict(stats::HoltWinters(fullts), n.ahead = 50, prediction.interval = 1))
pred$date = seq(as.Date("2018-01-01"), by = "month", length.out = nrow(pred))

plot(df[647:726,8], df[647:726, 4], type = "o", 
     main = "Predicted CO2 (Holt-Winters)", xlab = "Year", ylab = "CO2 (ppm)", 
     xlim = range(pred[,4], df[647:726,8]), ylim = range(pred[,2],df[647:726, 4] ))
matlines(pred[,4], pred[,-4], lwd = c(2,1,1), lty = c(1,2,2))
```
Prediction intervals for forecasting based on exponential smoothing methods require the forecast errors to be uncorrelated and normally distributed with mean zero and constant variance.

In contrast, Autoregressive Integrated Moving Average (ARIMA) models include an explicit statistical model for the irregular component of a time series, allowing non-zero autocorrelations in the irregular component. ARIMA models also require stationary time series.

Using ARIMA should be even more effective, because it includes the lag of each effect. By using a acf and pacf plot, we can see what lag seems reasonable, and feed this into our final SARIMA model. 
```{r}
residuals = fullts - season(fullts,12) - trend(fullts,12)
test(residuals)
auto.arima(residuals)

sarima = stats::arima(fullts, order = c(2,1,0), seasonal = list(order = c(2,1,1), period = 12))
sarimapred = data.frame(predict(sarima, n.ahead = 50, prediction.interval = 1))
sarimapred[,3:4] = cbind(sarimapred[,1] - 2*sarimapred[,2], sarimapred[,1] + 2*sarimapred[,2])
sarimapred$date = seq(as.Date("2018-01-01"), by = "month", length.out = nrow(sarimapred))

plot(df[647:726,8], df[647:726, 4], type = "o", main = "Predicted CO2 (SARIMA)",
     xlab = "Year", ylab = "CO2 (ppm)", xlim = range(sarimapred[,5], df[647:726,8]), 
     ylim = range(sarimapred[,4],df[647:726, 4] ))
matlines(sarimapred[,5], sarimapred[,c(1,3,4)], lwd = c(2,1,1), lty = c(1,2,2))

```
```{r}
library(GPFDA)
data(co2)
data_co2=co2
# store data into matrix and remove missing values
y=data.matrix(data_co2[,!names(data_co2)%in%'Annual_Average'])
data_co2[,!names(data_co2)%in%'Annual_Average']
y=matrix(t(y),ncol=1)
x=1:612/12; x[y<0]=NA
mat=cbind(y,x)
mat=na.omit(mat)
X=as.matrix(mat[,2])
Y=as.matrix(mat[,1])
x=as.matrix(seq(1,612,len=1000)/12)

?seq
```





```{r}
MAPE(pred[1:8,1], df[719:726,5])
MAPE(sarimapred[1:8,1], df[719:726,5])
```

